### 2019-06-05 Latex/Figures/_compactness.tex
draft revisions

% # •••••••••••••••••••••••••••••••••••••••••••••••••••••••••••••••••
% Compactness Example: District 6 of the Court Adopted Remedial Plan and District 7 of the 2011 Plan
% =================================================================
% -- FIGURE -- FIGURE -- FIGURE -- FIGURE -- FIGURE -- FIGURE --   % 
% ----------------------------------------------------------------- 
\begin{figure}
    \begin{center}
    \caption{Compactness Example: District 6 of the Court Adopted Remedial Plan and District 7 of the 2011 Plan}
    \label{fig:compact}
    \includegraphics[width=1\textwidth]{Figures/fig_compactness.pdf}
    \end{center}
    \tabnotes{District 7 of the Enacted plan is recognized as the ``Goofy kicking Donald Duck'', while District 6 of the Remedial map has the same geographic center.}
\end{figure}
% -----------------------------------------------------------------
% -- END FIGURE -- END FIGURE -- END FIGURE -- END FIGURE -- END FI %
% =================================================================