### 2019-06-05 Latex/Tables/_tab_congsum.tex

% =================================================================
% -- TABLE -- TABLE -- TABLE -- TABLE -- TABLE -- TABLE -- TABLE  % 
% -----------------------------------------------------------------
% US House Election Summaries (PA 2012-2016)
% ••••••••••••••••••••••••••••••••••••••••••••••••••••••••••••••••• 
%
\begin{table}[!htbp] \centering 
  \caption{US House Election Summaries (PA 2012-2016)} 
  \label{tab:congsum} 
\begin{tabular}{@{\extracolsep{-5pt}} ccccc} 
 & 2012 & 2014 & 2016 & AVE \\ 
\hline \\[-1.8ex] 
Seats &  [13R-5D] &  [13R-5D] &  [13R-5D] & [13R-5D] \\ 
Seat \% & 72.2\% & 72.2\% & 72.2\% & $0.720$ \\ 
Votes & 51.1\% & 55.5\% & 54.2\% & $0.533$ \\ 
Bias & 0.13 & 0.1 & 0.11 & $0.113$ \\ 
Efficiency Gap & 0.21 & 0.11 & 0.17 & $0.163$ \\ 
Mean/Median & 0.06 & 0.06 & 0.06 & $0.060$ \\ 
Declination & 0.46 & 0.36 & 0.39 & $0.403$ \\ 
\end{tabular}
\tabnotes{Uncontested races are imputed with 0.25 and 0.75 for the respective winners. All votes are calculated from the Republican perspective of the two-party vote. We've adjusted all gerrymandering measures such that negative numbers indicate bias in favor of the Democrats.}
\end{table} 
%
% -----------------------------------------------------------------
% -- END TABLE -- END TABLE -- END TABLE -- END TABLE -- END TABL %
% =================================================================